### 2016-11-29 local-capability-machine.tex
Added 'nsubsetsim' (almost like n-subset) and changed the standard islands.

%Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
\newcommand{\fun}{\rightarrow}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}
\newcommand{\blater}{\mathop{\blacktriangleright}}
\newcommand{\id}{\var{id}}
\newcommand{\undefined}{\mathit{undefined}}

@@

\begin{document}
\begin{flushright}
\today
\end{flushright}
\section{Domains and Notation}
\begin{align*}
\Addrs &::= \nats\\
\Words &::= \Caps + \ints \\
\Regs  &::= \RegName \rightarrow \Words\\
\Heaps &::= \Addrs \rightarrow \Words \\
\Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\}\\
\ExecConfs  &::= \Regs \times \Heaps \\
\Globals & ::= \{\glob, \local \} \\
\Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
\Confs &::= \ExecConfs + \{\failed \} + \halted \times \ExecConfs \\
\HeapSegments &::= \Addrs \parfun \Words
\end{align*}
Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capability is local or global. There are two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities. They are otherwise the same as their non-''permit write local'' counterparts.

@@

           [1 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma)]

@@

@@

  \item Show: $\iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.

@@

           SFTS $\iota_{\link} \nsubeq[n'-1] \iota^\pwl_{\link,\link+1} (\perma)$\\

@@

    & [4 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv} (\perma)]\\

@@

    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk} (\temp)]

@@

\iota^\nwl_{\start_\adv,\addrend_\adv} (\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv} (\perma)

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)

@@

Standard regions for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
  &\iota_{\start,\addrend}^\pwl \defeq \lambda v \ldotp
    \begin{cases}
      (v,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
      \bot & \text{otherwise}
    \end{cases} \\
  &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}
\lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
Revoking all temporary regions:
\begin{align*}
  \revokeTemp{} & : \Worlds \fun \Worlds \\
  \revokeTemp{W} & \defeq \lambda r \ldotp 
                   \begin{cases}
                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
                     W(r)                & \text{otherwise}
                   \end{cases}
\end{align*}
\newcommand{\wrev}[1]{\revokeTemp{#1}}

Standard regions for when write local is not allowed:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
    \begin{cases}
      (v,1,=,H^\nwl_{start,\addrend}) & v \neq \revoked \\
      \bot & \text{otherwise}
\end{cases}
\\ \\
  &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

@@

\begin{align*}
    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
    & \writeCond{}(\iota,\gl)(W) =  \\
    & \quad \begin{aligned}[t]
              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}

\begin{align*}
    & \readCond{}(\gl)(W) =  \\
    & \quad \begin{aligned}[t]
              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
                        & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
                        & \;\qquad W(r)\nsubeq[n-1] \iota^\pwl_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}

@@

Now define the value relation as follows:
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
           \end{aligned}
\end{align*}

@@

\subsubsection{Standard regions}
\begin{lemma}
  \begin{align*}
    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubeq \iota^\pwl_{\start,\addrend}(v)
  \end{align*}
\end{lemma}

>>>>

%Math notation
\newcommand{\restrictfun}[1]{|_{#1}}
\newcommand{\parfun}{\rightharpoonup}
\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
\newcommand{\fun}{\rightarrow}
\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}

\newcommand\subsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\subset$}\cr
  \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
\newcommand\supsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\supset$}\cr
  \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
\newcommand{\nsubsim}[1][n]{\stackrel{\tiny{#1}}{\subsetsim}}
\newcommand{\nsupsim}[1][n]{\stackrel{\tiny{#1}}{\supsetsim}}

\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}

\newcommand{\union}{\mathbin{\cup}}
\DeclareMathOperator{\dom}{dom}
\newcommand{\blater}{\mathop{\blacktriangleright}}
\newcommand{\id}{\var{id}}
\newcommand{\undefined}{\mathit{undefined}}

@@

\begin{document}
\begin{flushright}
\today
\end{flushright}
\section{Domains and Notation}

\begin{align*}
\Addrs &::= \nats\\
\Words &::= \Caps + \ints \\
\Regs  &::= \RegName \rightarrow \Words\\
\Heaps &::= \Addrs \rightarrow \Words \\
\Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\}\\
\ExecConfs  &::= \Regs \times \Heaps \\
\Globals & ::= \{\glob, \local \} \\
\Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
\Confs &::= \ExecConfs + \{\failed \} + \halted \times \ExecConfs \\
\HeapSegments &::= \Addrs \parfun \Words
\end{align*}
Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capability is local or global. There are two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities. They are otherwise the same as their non-''permit write local'' counterparts.

@@

           [1 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]

@@

\lau{Need to change this to use the new standard regions}

@@

  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$.

@@

           SFTS $\iota_{\link} \nsubsim[n'-1] \iota^\pwl_{\link,\link+1}$\\

@@

    & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\

@@

    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk}]

@@

\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}

@@

  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}

@@

Standard regions for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
\\
  &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}
\lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
Revoking all temporary regions:
\begin{align*}
  \revokeTemp{} & : \Worlds \fun \Worlds \\
  \revokeTemp{W} & \defeq \lambda r \ldotp 
                   \begin{cases}
                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
                     W(r)                & \text{otherwise}
                   \end{cases}
\end{align*}
\newcommand{\wrev}[1]{\revokeTemp{#1}}

Standard regions for when write local is not allowed:
\begin{align*}
  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,H^\nwl_{start,\addrend}) \\
\\
  &\iota_{\start,\addrend}^{\nwl,p} : \Regions \\
  &\iota_{\start,\addrend}^{\nwl,p} \defeq (\perma,1,=,H^\nwl_{start,\addrend})\\
\\
  &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

@@

We need a notion of subset between regions that is almost $n$-subset, but not quite. The only difference is that the view part of a region is diregarded. Define ``semi $n$-subset'' and ``semi $n$-supset'' as:
\lau{ Using normal the canonical $n$-subset gave some technical problems in the proofs. For instance, it was not possible to have a global read capability for a region governed by a permanent standard region. }
\begin{mathpar}
  \inferrule{(s,H) = (s',H) \\
              \forall \hat{W} \ldotp H \; s \; \hat{W} \nsubeq H' \; s' \; \hat{W} }
            { (v,s,\phi,H) \nsubsim (v',s',\phi',H')}
\end{mathpar}

\begin{align*}
    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
    & \writeCond{}(\iota,\gl)(W) =  \\
    & \quad \begin{aligned}[t]
              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \}
            \end{aligned}
\end{align*}

@@

Now define the value relation as follows:
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
           \end{aligned}
\end{align*}

@@

\subsubsection{Standard regions}
\begin{lemma}
  \begin{align*}
    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
  \end{align*}
\end{lemma}
